Some notes for developing the code.\subsection{Coding Style}\label{devel_codingstyle}
\begin{itemize}
\item We try to follow ANSI C standard as close as possible, however we use also some advantages of C99. Thus, you should use ANSI C where possible, and C99 where necessary.\item We use ANSI C comment style. Because of Doxygen documenting style, use one asteriks {\tt $\ast$} at start of your non-documentation comments.\item We use 2 spaces indenting style (you can easily map tabs to 2 spaces).\item Try not to use globals. Avoid them where possible.\item Remember, the code should be readable by humans. For the machines, it does not matter.\item We follow {\tt PEAR} coding standards, see {\tt http://pear.php.net/manual/en/standards.php}\end{itemize}


If you are a lucky vim user, try settings below: 

\begin{Code}\begin{verbatim} :set textwidth=79
 :set shiftwidth=2
 :set tabstop=2
 :set smarttab
 :set expandtab
 :set list
 :highlight OverLength ctermbg=red ctermfg=white guibg=#592929
 :match OverLength /\%80v.* /
 :let c_space_errors = 1
\end{verbatim}
\end{Code}

\subsection{Message Interface}\label{devel_message}
{\tt Mechanic} provides some functions that should be used instead of standard {\tt printf} and {\tt exit:} 

\begin{itemize}
\item 

\begin{Code}\begin{verbatim} void mechanic_message(int type, char* fmt, ...) 
\end{verbatim}
\end{Code}

 A wrapper for any message printing. The available types are:\begin{itemize}
\item {\tt MECHANIC\_\-MESSAGE\_\-INFO} -- information style\item {\tt MECHANIC\_\-MESSAGE\_\-CONT} -- continuation of any message style (when the message is too long, and you want to split it)\item {\tt MECHANIC\_\-MESSAGE\_\-ERR} -- error only\item {\tt MECHANIC\_\-MESSAGE\_\-WARN} -- warning only\item {\tt MECHANIC\_\-MESSAGE\_\-DEBUG} -- debug only\end{itemize}
\end{itemize}


The {\tt $\ast$fmt} is a standard {\tt printf} format, and dots {\tt }... are arguments for it.

\begin{itemize}
\item 

\begin{Code}\begin{verbatim} mechanic_error(int stat) 
\end{verbatim}
\end{Code}

 Handles error states, see \doxyref{Error Codes}{p.}{troubleshooting_errcodes} for available error codes.\end{itemize}


\begin{itemize}
\item 

\begin{Code}\begin{verbatim} mechanic_abort(int node) 
\end{verbatim}
\end{Code}

 A wrapper for {\tt MPI\_\-Abort}.\end{itemize}


\begin{itemize}
\item 

\begin{Code}\begin{verbatim} mechanic_finalize(int node) 
\end{verbatim}
\end{Code}

 A wrapper for {\tt MPI\_\-Finalize}. \end{itemize}
